```{r setoptions, echo=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```

Peer assessment 1
==============================

## Loading and preprocessing the data
```{r}
data <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
```{r}
sum_per_day <- aggregate(data$steps, by=list(data$date), FUN=sum, na.rm=T)
hist(sum_per_day$x, main="Total steps per day", xlab="Total steps")
```

```{r}
mean_total <- mean(sum_per_day$x)
median_total <- median(sum_per_day$x)
```
The mean of total steps per day is `r mean_total`.
The median of total steps per day is `r median_total`.


## What is the average daily activity pattern?
```{r}
mean_per_interval <- aggregate(data$steps, by=list(data$interval), FUN=mean, na.rm=T)
plot(mean_per_interval, type="l", xlab="Interval", ylab="Steps")
max_interval <- mean_per_interval[which.max(mean_per_interval$x), 1]
```
The interval which contains the maximum number of steps across all the days on average is `r max_interval`.


## Imputing missing values
```{r}
total_na <- sum(is.na(data))

data2 <- data.frame(data)
for(i in 1:nrow(data)){
    if (is.na(data[i,])[1]){
        int <- data[i,3]
        data2[i,1] = mean_per_interval[mean_per_interval[,1] == int, 2]
    }
}

sum_per_day2 <- aggregate(data2$steps, by=list(data2$date), FUN=sum, na.rm=T)
hist(sum_per_day2$x, main="Total steps per day", xlab="Total steps")
```
Total NA lines: `r total_na`.
After the substitution, the distribution became "more normal". That is not surprising because we have changed the data in the most expected way for each missed value (we used the mean values for the corresponding intervals instead NA).

## Are there differences in activity patterns between weekdays and weekends?
Adding the column to data2:
```{r results="hide"}
x <- rep("weekday", nrow(data))
x[which(weekdays(as.Date(data2$date)) %in% c("Saturday", "Sunday"))] = "weekend"
x <- as.factor(x)
data2 <- cbind(data2, x)
colnames(data2)[4] <- "weekend"
```

Adding a temporary data frame `mean_per_interval` with aggregated average values
```{r results="hide"}
mean_per_interval <- aggregate(data2$steps, by=list(data2$interval, data2$weekend), FUN=mean, na.rm=T)
colnames(mean_per_interval) <- c("interval", "weekend", "steps")
```

Plotting the results.
```{r}
library(lattice)
with(mean_per_interval,
    xyplot(steps ~ interval | weekend, type="l", layout=c(1,2))
)
```